Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-10 paper.qmd
Updated Figure 11

trimesh_removed_gau_tsne <- ggplot() + 
  geom_point(
    data = tsne_gau_scaled,
    aes(
      x = tSNE1,
      y = tSNE2
    ),
    alpha = 0.2) + 
  # geom_rect(
  #     aes(xmin = 0.3, 
  #         xmax = 0.65, 
  #         ymin = -0.05, 
  #         ymax = 0.35),
  #         fill = "transparent", 
  #     color = "grey50", 
  #     linewidth = 0.5) +
  interior_annotation("a1", 
                      position = c(0.08, 0.95),
                      cex = 1.5) +
  theme(
    aspect.ratio = 1
  )

@@

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.3,
    color = "#999999") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.1) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2),
    label=rownames(axes), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 1.5)
```

@@

trimesh_removed_gau_umap <- ggplot() + 
  geom_point(
    data = umap_gau_scaled,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    alpha = 0.05) + 
  # geom_rect(
  #   aes(xmin = 0.45, 
  #       xmax = 0.57, 
  #       ymin = 0, 
  #       ymax = 0.1),
  #   fill = "transparent", 
  #   color = "grey50", 
  #   linewidth = 0.5)  +
  interior_annotation("b1", 
                      position = c(0.06, 0.95),
                      cex = 1.5) +
  theme(
    aspect.ratio = 1
  )

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.3,
    color = "#999999") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.1) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("b2", 
                      position = c(0.08, 0.9),
                      cex = 1.5)
```

@@

trimesh_removed_gau_pacmap <- ggplot() + 
  geom_point(
    data = pacmap_gau_scaled,
    aes(
      x = PaCMAP1,
      y = PaCMAP2
    ),
    alpha = 0.3) +
  # geom_rect(
  #   aes(xmin = -0.04, 
  #       xmax = 0.22, 
  #       ymin = 0.08, 
  #       ymax = 0.34),
  #   fill = "transparent", 
  #   color = "grey50", 
  #   linewidth = 0.5) +
  interior_annotation("c1", 
                      position = c(0.08, 0.95),
                      cex = 1.5) +
  theme(
    aspect.ratio = 1
  )

@@

five_gau_proj_pacmap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1,
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.3,
    color = "#999999") +
  geom_segment(
    data = model_df,
    aes(
      x = proj1_from,
      y = proj2_from,
      xend = proj1_to,
      yend = proj2_to),
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.1) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2, label=rownames(axes)),
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle,
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("c2",
                      position = c(0.08, 0.9),
                      cex = 1.5)

>>>>

trimesh_removed_gau_tsne <- ggplot() + 
  geom_point(
    data = tsne_gau_scaled,
    aes(
      x = tSNE1,
      y = tSNE2
    ),
    alpha = 0.2) + 
  # geom_rect(
  #     aes(xmin = 0.3, 
  #         xmax = 0.65, 
  #         ymin = -0.05, 
  #         ymax = 0.35),
  #         fill = "transparent", 
  #     color = "grey50", 
  #     linewidth = 0.5) +
  interior_annotation("a1", 
                      position = c(0.08, 0.95),
                      cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.3,
    color = "#999999") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.1) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2),
    label=rownames(axes), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 2)
```

@@

trimesh_removed_gau_umap <- ggplot() + 
  geom_point(
    data = umap_gau_scaled,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    alpha = 0.05) + 
  # geom_rect(
  #   aes(xmin = 0.45, 
  #       xmax = 0.57, 
  #       ymin = 0, 
  #       ymax = 0.1),
  #   fill = "transparent", 
  #   color = "grey50", 
  #   linewidth = 0.5)  +
  interior_annotation("b1", 
                      position = c(0.95, 0.95),
                      cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.3,
    color = "#999999") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.1) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("b2", 
                      position = c(0.08, 0.9),
                      cex = 2)
```

@@

trimesh_removed_gau_pacmap <- ggplot() + 
  geom_point(
    data = pacmap_gau_scaled,
    aes(
      x = PaCMAP1,
      y = PaCMAP2
    ),
    alpha = 0.3) +
  # geom_rect(
  #   aes(xmin = -0.04, 
  #       xmax = 0.22, 
  #       ymin = 0.08, 
  #       ymax = 0.34),
  #   fill = "transparent", 
  #   color = "grey50", 
  #   linewidth = 0.5) +
  interior_annotation("c1", 
                      position = c(0.08, 0.95),
                      cex = 2) +
  theme(
    aspect.ratio = 1
  )

@@

five_gau_proj_pacmap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1,
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.3,
    color = "#999999") +
  geom_segment(
    data = model_df,
    aes(
      x = proj1_from,
      y = proj2_from,
      xend = proj1_to,
      yend = proj2_to),
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.1) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2, label=rownames(axes)),
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle,
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("c2",
                      position = c(0.08, 0.9),
                      cex = 2)
